Shrinking Characteristics of Precision Matrix Estimators: An Illustration via Regression

### 2019-01-09 02-PME.Rmd
restructed manuscript

# `ADMMsigma` Details

@@

\begin{align}
  f(X_{1}, ..., X_{n}; \mu, \Sigma) &= \prod_{i = 1}^{n}(2\pi)^{-p/2}\left| \Sigma \right|^{-1/2}\exp\left[ -\frac{1}{2}\left( X_{i} - \mu \right)^{T}\Sigma^{-1}\left( X_{i} - \mu \right) \right] \notag \\
  &= (2\pi)^{-nr/2}\left| \Sigma \right|^{-n/2}\mbox{etr}\left[ -\frac{1}{2}\sum_{i = 1}^{n}\left( X_{i} - \mu \right)\left( X_{i} - \mu \right)^{T}\Sigma^{-1} \right]
\end{align}

@@

\begin{equation}
  l(\mu, \Sigma | X) = const. - \frac{n}{2}\log\left| \Sigma \right| - tr\left[ \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)^{T}\Sigma^{-1} \right]
(\#eq:loglik)
\end{equation}

@@

\begin{align*}
  &\hat{\Sigma}_{MLE} = \arg\max_{\Sigma \in \mathbb{S}_{+}^{p}}\left\{ const. - \frac{n}{2}\log\left| \Sigma \right| - tr\left[ \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)^{T}\Sigma^{-1} \right] \right\} \\
  &\nabla_{\Sigma}l(\mu, \Sigma | X) = -\frac{n}{2}\Sigma^{-1} + \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)^{T}\Sigma^{-2} \\
  \Rightarrow &\hat{\Sigma}_{MLE} = \left[ \frac{1}{n}\sum_{i = 1}^{n}\left(X_{i} - \bar{X} \right)\left(X_{i} - \bar{X} \right)^{T} \right]
\end{align*}

@@

\[ \hat{\Omega}_{MLE} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| \right\} \]

@@

\[ \hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + P\left( \Omega \right) \right\} \]

@@

\[ P\left( \Omega \right) = \lambda\left[\frac{1 - \alpha}{2}\left\| \Omega \right|_{F}^{2} + \alpha\left\| \Omega \right\|_{1} \right] \]

@@

\[ \hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + \lambda\left[\frac{1 - \alpha}{2}\left\| \Omega \right|_{F}^{2} + \alpha\left\| \Omega \right\|_{1} \right] \right\} \]

@@

@@

<br>\vspace{1cm}

## ADMM Algorithm

@@

\begin{align*}
  \mbox{minimize } f(x) + g(z) \\
  \mbox{subject to } Ax + Bz = c
\end{align*}

@@

\[ L_{\rho}(x, z, y) = f(x) + g(z) + y^{T}(Ax + Bz - c) + \frac{\rho}{2}\left\| Ax + Bz - c \right\|_{2}^{2} \]

@@

\begin{align}
  x^{k + 1} &= \arg\min_{x \in \mathbb{R}^{n}}L_{\rho}(x, z^{k}, y^{k}) \\
  z^{k + 1} &= \arg\min_{z \in \mathbb{R}^{m}}L_{\rho}(x^{k + 1}, z, y^{k}) \\
  y^{k + 1} &= y^{k} + \rho(Ax^{k + 1} + Bz^{k + 1} - c)
\end{align}

@@

\[ L_{\rho}(\Omega, Z, \Lambda) = f\left(\Omega\right) + g\left(Z\right) + tr\left[\Lambda\left(\Omega - Z\right)\right] \]

@@

\[ L_{\rho}(\Omega, Z, \Lambda) = f\left(\Omega\right) + g\left(Z\right) + tr\left[\Lambda\left(\Omega - Z\right)\right] + \frac{\rho}{2}\left\|\Omega - Z\right\|_{F}^{2} \]

@@

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \\
  Z^{k + 1} &= \arg\min_{Z \in \mathbb{S}^{p}}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \\
  \Lambda^{k + 1} &= \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right)
\end{align}

<br>\vspace{1cm}

@@

1. Decompose $S + \Lambda^{k} - \rho Z^{k} = VQV^{T}$ via spectral decomposition.

\[ \Omega^{k + 1} = \frac{1}{2\rho}V\left[ -Q + \left( Q^{2} + 4\rho I_{p} \right)^{1/2} \right]V^{T} \]

@@

\begin{align*}
Z_{ij}^{k + 1} &= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{sign}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}\right)\left( \left| \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} \right| - \lambda\alpha \right)_{+} \\
&= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{soft}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}, \lambda\alpha\right)
\end{align*}

@@

\[ \Lambda^{k + 1} = \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right) \]

@@

<br>\vspace{1cm}

### Proof of (1):

\[ \Omega^{k + 1} = \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \]

\vspace{1cm}

\begin{align*}
  \nabla_{\Omega}&\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \\
  &= S - \Omega^{-1} + \Lambda^{k} + \rho\left( \Omega - Z^{k} \right)
\end{align*}

Note that because all of the variables are symmetric, we can ignore the symmetric constraint when deriving the gradient. First set the gradient equal to zero and decompose $\Omega^{k + 1} = VDV^{T}$ where $D$ is a diagonal matrix with diagonal elements equal to the eigen values of $\Omega^{k + 1}$ and $V$ is the matrix with corresponding eigen vectors as columns:

\[ S + \Lambda^{k} - \rho Z^{k} = (\Omega^{k + 1})^{-1} - \rho \Omega^{k + 1} = VD^{-1}V^{T} - \rho VDV^{T} =  V\left(D^{-1} - \rho D\right)V^{T} \]

This equivalence implies that

\[ \phi_{j}\left( S + \Lambda^{k} - \rho Z^{k} \right) = \frac{1}{\phi_{j}(\Omega^{k + 1})} - \rho\phi_{j}(\Omega^{k + 1}) \]

where $\phi_{j}(\cdot)$ is the $j$th eigen value.

\begin{align*}
  &\Rightarrow \rho\phi_{j}^{2}(\Omega^{k + 1}) + \phi_{j}\left( S + \Lambda^{k} - \rho Z^{k} \right)\phi_{j}(\Omega^{k + 1}) - 1 = 0 \\
  &\Rightarrow \phi_{j}(\Omega^{k + 1}) = \frac{-\phi_{j}(S + \Lambda^{k} - \rho Z^{k}) \pm \sqrt{\phi_{j}^{2}(S + \Lambda^{k} - \rho Z^{k}) + 4\rho}}{2\rho}
\end{align*}

In summary, if we decompose $S + \Lambda^{k} - \rho Z^{k} = VQV^{T}$ then

\[ \Omega^{k + 1} = \frac{1}{2\rho}V\left[ -Q + (Q^{2} + 4\rho I_{p})^{1/2}\right] V^{T} \]

<br>\vspace{1cm}

### Proof of (2)

\[ Z^{k + 1} = \arg\min_{Z \in \mathbb{S}^{p}}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \]

\vspace{1cm}

\begin{align*}
  \partial&\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \\
  &= \partial\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] \right\} + \nabla_{\Omega}\left\{ tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \\
  &= \lambda(1 - \alpha)Z + \mbox{sign}(Z)\lambda\alpha - \Lambda^{k} - \rho\left( \Omega^{k + 1} - Z \right)
\end{align*}

where sign is the elementwise sign operator. By setting the gradient/sub-differential equal to zero, we arrive at the following equivalence:

\[ Z_{ij}^{k + 1} = \frac{1}{\lambda(1 - \alpha) + \rho}\left( \rho \Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} - \mbox{sign}(Z_{ij}^{k + 1})\lambda\alpha \right) \]

for all $i = 1,..., p$ and $j = 1,..., p$. We observe two scenarios:

- If $Z_{ij}^{k + 1} > 0$ then

\[ \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} > \lambda\alpha \]

- If $Z_{ij}^{k + 1} < 0$ then

\[ \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} < -\lambda\alpha \]

This implies that $\mbox{sign}(Z_{ij}) = \mbox{sign}(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k})$. Putting all the pieces together, we arrive at

\begin{align*}
Z_{ij}^{k + 1} &= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{sign}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}\right)\left( \left| \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} \right| - \lambda\alpha \right)_{+} \\
&= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{soft}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}, \lambda\alpha\right)
\end{align*}

where soft is the soft-thresholding function.

<br>\vspace{1cm}

## Scaled-Form ADMM

@@

\begin{align*}
  tr\left[ \Lambda^{k}\left( \Omega - Z^{k} \right) \right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} &= tr\left[ \Lambda^{k}R^{k} \right] + \frac{\rho}{2}\left\| R^{k} \right\|_{F}^{2} \\
  &= \frac{\rho}{2}\left\| R^{k} + \Lambda^{k}/\rho \right\|_{F}^{2} - \frac{\rho}{2}\left\| \Lambda^{k}/\rho \right\|_{F}^{2} \\
  &= \frac{\rho}{2}\left\| R^{k} + U^{k} \right\|_{F}^{2} - \frac{\rho}{2}\left\| U^{k} \right\|_{F}^{2}
\end{align*}

@@

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega \in \mathbb{R}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + \frac{\rho}{2}\left\| \Omega - Z^{k} + U^{k} \right\|_{F}^{2} \right\} \\
  Z^{k + 1} &= \arg\min_{Z \in \mathbb{S}^{p}}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z + U^{k} \right\|_{F}^{2} \right\} \\
  U^{k + 1} &= U^{k} + \Omega^{k + 1} - Z^{k + 1}
\end{align}

@@

\begin{align}
  x^{k + 1} &= \arg\min_{x}\left\{ f(x) + \frac{\rho}{2}\left\| Ax + Bz^{k} - c + u^{k} \right\|_{2}^{2} \right\} \\
  z^{k + 1} &= \arg\min_{z}\left\{ g(z) + \frac{\rho}{2}\left\| Ax^{k + 1} + Bz - c + u^{k} \right\|_{2}^{2} \right\} \\
  u^{k + 1} &= u^{k} + Ax^{k + 1} + Bz^{k + 1} - c
\end{align}

@@

<br>\vspace{1cm}

## Stopping Criterion

@@

\[ \Omega^{k + 1} - Z^{k + 1} = 0 \]

@@

\[ 0 \in \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k + 1} \]

\[ 0 \in \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1} \]

@@

\begin{align*}
  0 &\in \partial \left\{ f\left(\Omega^{k + 1}\right) + tr\left[ \Lambda^{k}\left( \Omega^{k + 1} - Z^{k} \right) \right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z^{k} \right\|_{F}^{2} \right\} \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k} + \rho\left(\Omega^{k + 1} - Z^{k}\right) \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k} + \rho\left(\Omega^{k + 1} + Z^{k + 1} - Z^{k + 1} - Z^{k}\right) \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k} + \rho\left(\Omega^{k + 1} - Z^{k + 1}\right) + \rho\left(Z^{k + 1} - Z^{k}\right) \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k + 1} + \rho\left(Z^{k + 1} - Z^{k}\right) \\
  \Rightarrow 0 &\in \rho\left( Z^{k + 1} - Z^{k} \right)
\end{align*}

@@

\begin{align*}
  0 &\in \partial \left\{ g\left(Z^{k + 1}\right) + tr\left[ \Lambda^{k}\left( \Omega^{k + 1} - Z^{k + 1} \right) \right] + \rho\left\| \Omega^{k + 1} - Z^{k + 1} \right\|_{F}^{2} \right\} \\
  &= \partial g\left(Z^{k + 1}\right) - \Lambda^{k} - \rho\left(\Omega^{k + 1} - Z^{k + 1}\right) \\
  &= \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1} \\
\end{align*}

@@

\begin{align*}
  \epsilon^{pri} &= p\epsilon^{abs} + \epsilon^{rel}\max\left\{ \left\| \Omega^{k + 1} \right\|_{F}, \left\| Z^{k + 1} \right\|_{F} \right\} \\
  \epsilon^{dual} &= p\epsilon^{abs} + \epsilon^{rel}\left\| \Lambda^{k + 1} \right\|_{F}
\end{align*}

>>>>

# Precision Matrix Estimation

## Background

@@

\begin{equation} 
\begin{split}
  f(X_{1}, ..., X_{n}; \mu, \Sigma) &= \prod_{i = 1}^{n}(2\pi)^{-p/2}\left| \Sigma \right|^{-1/2}\exp\left[ -\frac{1}{2}\left( X_{i} - \mu \right)'\Sigma^{-1}\left( X_{i} - \mu \right) \right] \\
  &= (2\pi)^{-nr/2}\left| \Sigma \right|^{-n/2}\mbox{etr}\left[ -\frac{1}{2}\sum_{i = 1}^{n}\left( X_{i} - \mu \right)\left( X_{i} - \mu \right)'\Sigma^{-1} \right]
\end{split}
(\#eq:lik)
\end{equation}

@@

\begin{equation}
  l(\mu, \Sigma | X) = const. - \frac{n}{2}\log\left| \Sigma \right| - tr\left[ \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)'\Sigma^{-1} \right]
(\#eq:loglik)
\end{equation}

@@

\begin{equation}
\hat{\Sigma}^{mle} = \arg\max_{\Sigma \in \mathbb{S}_{+}^{p}}\left\{ const. - \frac{n}{2}\log\left| \Sigma \right| - tr\left[ \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)'\Sigma^{-1} \right] \right\}
(\#eq:sigmaloglik)
\end{equation}

\begin{equation}
\nabla_{\Sigma}l(\mu, \Sigma | X) = -\frac{n}{2}\Sigma^{-1} + \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)'\Sigma^{-2}
(\#eq:nablasigma)
\end{equation}

\begin{equation}
\Rightarrow \hat{\Sigma}^{mle} = \left[ \frac{1}{n}\sum_{i = 1}^{n}\left(X_{i} - \bar{X} \right)\left(X_{i} - \bar{X} \right)' \right]
(\#eq:sigmamle)
\end{equation}

@@

\begin{equation}
\hat{\Omega}^{mle} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| \right\}
(\#eq:omegamle)
\end{equation}

@@

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + P\left( \Omega \right) \right\}
(\#eq:omegapen)
\end{equation}

@@

\begin{equation}
P\left( \Omega \right) = \lambda\left[\frac{1 - \alpha}{2}\left\| \Omega \right|_{F}^{2} + \alpha\left\| \Omega \right\|_{1} \right]
(\#eq:pen)
\end{equation}

@@

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + \lambda\left[\frac{1 - \alpha}{2}\left\| \Omega \right|_{F}^{2} + \alpha\left\| \Omega \right\|_{1} \right] \right\}
(\#eq:omegaloglik)
\end{equation}

@@

The penalty proposed by @molstad2017shrinking is one of the following form:

\begin{equation}
P\left(\Omega\right) = \lambda\left\| A\Omega B - C \right\|_{1}
(\#eq:pen2)
\end{equation}

where $A \in \mathbb{R}^{m \times p}, B \in \mathbb{R}^{p \times q}, \mbox{ and } C \in \mathbb{R}^{m \times q}$ are matrices assumed to be known and specified by the user. Solving the full penalized log-likelihood for $\Omega$ results in solving

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + \lambda\left\| A\Omega B - C \right\|_{1} \right\}
(\#eq:omegaloglik2)
\end{equation}

This form of penalty is particularly useful because matrices $A, B, \mbox{ and } C$ can be constructed so that we penalize the sum, absolute value of a *characteristic* of the precision matrix $\Omega$. This type of penalty leads to many new, interesting, and novel estimators for $\Omega$. An example of one such estimator (suppose we observe $n$ samples of $Y_{i} \in \mathbb{R}^{r}$) would be one where we set $A = I_{p}, B = \Sigma_{xy}, \mbox{ and } C = 0$ where $\Sigma_{xy}$ is the covariance matrix of $X$ and $Y$. This penalty has the effect of assuming sparsity in the forward regression coefficient $\beta \equiv \Omega\Sigma_{xy}$. Of course, in practice we do not know the true covariance matrix $\Sigma_{xy}$ but we might consider using the sample estimate $\hat{\Sigma}_{xy} = \sum_{i = 1}^{n}\left(X_{i} - \bar{X}\right)\left(Y_{i} - \bar{Y}\right)^{T}/n$

@@

## ADMM algorithm

@@

\begin{equation}
\begin{split}
  \mbox{minimize } f(x) + g(z) \\
  \mbox{subject to } Ax + Bz = c
\end{split}
\end{equation}

@@

\begin{equation}
L_{\rho}(x, z, y) = f(x) + g(z) + y'(Ax + Bz - c) + \frac{\rho}{2}\left\| Ax + Bz - c \right\|_{2}^{2}
(\#eq:auglagrange)
\end{equation}

@@

\begin{align}
  x^{k + 1} &= \arg\min_{x \in \mathbb{R}^{n}}L_{\rho}(x, z^{k}, y^{k}) (\#eq:xauglagrange) \\
  z^{k + 1} &= \arg\min_{z \in \mathbb{R}^{m}}L_{\rho}(x^{k + 1}, z, y^{k}) (\#eq:zauglagrange) \\
  y^{k + 1} &= y^{k} + \rho(Ax^{k + 1} + Bz^{k + 1} - c) (\#eq:yauglagrange)
\end{align}

@@

\begin{equation}
L_{\rho}(\Omega, Z, \Lambda) = f\left(\Omega\right) + g\left(Z\right) + tr\left[\Lambda\left(\Omega - Z\right)\right]
(\#eq:lagrange)
\end{equation}

@@

\begin{equation}
L_{\rho}(\Omega, Z, \Lambda) = f\left(\Omega\right) + g\left(Z\right) + tr\left[\Lambda\left(\Omega - Z\right)\right] + \frac{\rho}{2}\left\|\Omega - Z\right\|_{F}^{2}
(\#eq:auglagrangian)
\end{equation}

@@

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} (\#eq:omegaauglagrange) \\
  Z^{k + 1} &= \arg\min_{Z \in \mathbb{S}^{p}}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} (\#eq:ZZauglagrange) \\
  \Lambda^{k + 1} &= \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right) (\#eq:lamauglagrange)
\end{align}

@@

1. Decompose $S + \Lambda^{k} - \rho Z^{k} = VQV'$ via spectral decomposition.

\begin{equation}
\Omega^{k + 1} = \frac{1}{2\rho}V\left[ -Q + \left( Q^{2} + 4\rho I_{p} \right)^{1/2} \right]V'
(\#eq:omegaalgo)
\end{equation}

@@

\begin{equation}
\begin{split}
Z_{ij}^{k + 1} &= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{sign}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}\right)\left( \left| \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} \right| - \lambda\alpha \right)_{+} \\
&= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{soft}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}, \lambda\alpha\right)
(\#eq:ZZalgo)
\end{split}
\end{equation}

@@

\begin{equation}
\Lambda^{k + 1} = \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right)
(\#eq:lamalgo)
\end{equation}

@@

### Scaled-Form ADMM

@@

\begin{equation}
\begin{split}
  tr\left[ \Lambda^{k}\left( \Omega - Z^{k} \right) \right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} &= tr\left[ \Lambda^{k}R^{k} \right] + \frac{\rho}{2}\left\| R^{k} \right\|_{F}^{2} \\
  &= \frac{\rho}{2}\left\| R^{k} + \Lambda^{k}/\rho \right\|_{F}^{2} - \frac{\rho}{2}\left\| \Lambda^{k}/\rho \right\|_{F}^{2} \\
  &= \frac{\rho}{2}\left\| R^{k} + U^{k} \right\|_{F}^{2} - \frac{\rho}{2}\left\| U^{k} \right\|_{F}^{2}
(\#eq:scaled)
\end{split}
\end{equation}

@@

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega \in \mathbb{R}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + \frac{\rho}{2}\left\| \Omega - Z^{k} + U^{k} \right\|_{F}^{2} \right\} (\#eq:omegascaled) \\
  Z^{k + 1} &= \arg\min_{Z \in \mathbb{S}^{p}}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z + U^{k} \right\|_{F}^{2} \right\} (\#eq:ZZscaled) \\
  U^{k + 1} &= U^{k} + \Omega^{k + 1} - Z^{k + 1} (\#eq:Uscaled)
\end{align}

@@

\begin{align}
  x^{k + 1} &= \arg\min_{x}\left\{ f(x) + \frac{\rho}{2}\left\| Ax + Bz^{k} - c + u^{k} \right\|_{2}^{2} \right\} (\#eq:xscaled) \\
  z^{k + 1} &= \arg\min_{z}\left\{ g(z) + \frac{\rho}{2}\left\| Ax^{k + 1} + Bz - c + u^{k} \right\|_{2}^{2} \right\} (\#eq:zscaled) \\
  u^{k + 1} &= u^{k} + Ax^{k + 1} + Bz^{k + 1} - c (\#eq:uscaled)
\end{align}

@@

### Stopping Criterion

@@

\begin{equation}
\Omega^{k + 1} - Z^{k + 1} = 0
(\#eq:admmprimal)
\end{equation}

@@

\begin{equation}
\begin{split}
  0 &\in \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k + 1} \\
  0 &\in \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1}
(\#eq:admmdual)
\end{split}
\end{equation}

@@

\begin{equation}
\begin{split}
  0 &\in \partial \left\{ f\left(\Omega^{k + 1}\right) + tr\left[ \Lambda^{k}\left( \Omega^{k + 1} - Z^{k} \right) \right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z^{k} \right\|_{F}^{2} \right\} \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k} + \rho\left(\Omega^{k + 1} - Z^{k}\right) \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k} + \rho\left(\Omega^{k + 1} + Z^{k + 1} - Z^{k + 1} - Z^{k}\right) \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k} + \rho\left(\Omega^{k + 1} - Z^{k + 1}\right) + \rho\left(Z^{k + 1} - Z^{k}\right) \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k + 1} + \rho\left(Z^{k + 1} - Z^{k}\right) \\
  \Rightarrow 0 &\in \rho\left( Z^{k + 1} - Z^{k} \right)
(\#eq:dualresidual)
\end{split}
\end{equation}

@@

\begin{equation}
\begin{split}
  0 &\in \partial \left\{ g\left(Z^{k + 1}\right) + tr\left[ \Lambda^{k}\left( \Omega^{k + 1} - Z^{k + 1} \right) \right] + \rho\left\| \Omega^{k + 1} - Z^{k + 1} \right\|_{F}^{2} \right\} \\
  &= \partial g\left(Z^{k + 1}\right) - \Lambda^{k} - \rho\left(\Omega^{k + 1} - Z^{k + 1}\right) \\
  &= \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1} \\
(\#eq:dualopt)
\end{split}
\end{equation}

@@

\begin{equation}
\begin{split}
  \epsilon^{pri} &= p\epsilon^{abs} + \epsilon^{rel}\max\left\{ \left\| \Omega^{k + 1} \right\|_{F}, \left\| Z^{k + 1} \right\|_{F} \right\} \\
  \epsilon^{dual} &= p\epsilon^{abs} + \epsilon^{rel}\left\| \Lambda^{k + 1} \right\|_{F}
(\#eq:admmstopping)
\end{split}
\end{equation}

## Simulations

In the simulations below, we generate data from a number of different oracle precision matrices with various structures. For each data-generating procedure, the `ADMMsigma()` function was run using 5-fold cross validation. After 20 replications, the cross validation errors were totalled and the optimal tuning parameters were selected (results below the top figure). These results are compared with the Kullback Leibler (KL) losses between the estimates and the oracle precision matrix (bottom figure).

We can see below that our cross validation procedure consistently chooses tuning parameters that are close to the optimal parameters with respsect to the oracle.

### Plots

#### Compound Symmetric: P = 100, N = 50

```{r, message = FALSE, echo = TRUE, eval = FALSE}

# oracle precision matrix
Omega = matrix(0.9, ncol = 100, nrow = 100)
diag(Omega = 1)

# generate covariance matrix
S = qr.solve(Omega)

# generate data
Z = matrix(rnorm(100*50), nrow = 50, ncol = 100)
out = eigen(S, symmetric = TRUE)
S.sqrt = out$vectors %*% diag(out$values^0.5) %*% t(out$vectors)
X = Z %*% S.sqrt

```
<br>\vspace{0.5cm}

<!-- <center> -->
<!-- ![Figure caption here.](images/compound_N50_P100.png){ -->
<!-- width=90% } -->
<!-- </center> -->

```{r admmsim1, eval = T, echo = F, fig.cap = "Figure caption here."}

knitr::include_graphics("images/compound_N50_P100.png")

```

#### Compound Symmetric: P = 10, N = 1000

```{r, message = FALSE, echo = TRUE, eval = FALSE}

# oracle precision matrix
Omega = matrix(0.9, ncol = 10, nrow = 10)
diag(Omega = 1)

# generate covariance matrix
S = qr.solve(Omega)

# generate data
Z = matrix(rnorm(10*1000), nrow = 1000, ncol = 10)
out = eigen(S, symmetric = TRUE)
S.sqrt = out$vectors %*% diag(out$values^0.5) %*% t(out$vectors)
X = Z %*% S.sqrt

```
<br>\vspace{0.5cm}

<!-- <center> -->
<!-- ![Figure caption here.](images/compound_N1000_P10.png){ -->
<!-- width=90% } -->
<!-- </center> -->

```{r admmsim2, eval = T, echo = F, fig.cap = "Figure caption here."}

knitr::include_graphics("images/compound_N1000_P10.png")

```

#### Dense: P = 100, N = 50

```{r, message = FALSE, echo = TRUE, eval = FALSE, tidy = FALSE}

# generate eigen values
eigen = c(rep(1000, 5, rep(1, 100 - 5)))

# randomly generate orthogonal basis (via QR)
Q = matrix(rnorm(100*100), nrow = 100, ncol = 100) %>% qr %>% qr.Q

# generate covariance matrix
S = Q %*% diag(eigen) %*% t(Q)

# generate data
Z = matrix(rnorm(100*50), nrow = 50, ncol = 100)
out = eigen(S, symmetric = TRUE)
S.sqrt = out$vectors %*% diag(out$values^0.5) %*% t(out$vectors)
X = Z %*% S.sqrt

```
<br>\vspace{0.5cm}

<!-- <center> -->
<!-- ![Figure caption here.](images/repsKLdenseQR_N50_P100.png){ -->
<!-- width=90% } -->
<!-- </center> -->
<!-- } -->

```{r admmsim3, eval = T, echo = F, fig.cap = "Figure caption here."}

knitr::include_graphics("images/repsKLdenseQR_N50_P100.png")

```

#### Dense: P = 10, N = 50

```{r, message = FALSE, echo = TRUE, eval = FALSE, tidy = FALSE}

# generate eigen values
eigen = c(rep(1000, 5, rep(1, 10 - 5)))

# randomly generate orthogonal basis (via QR)
Q = matrix(rnorm(10*10), nrow = 10, ncol = 10) %>% qr %>% qr.Q

# generate covariance matrix
S = Q %*% diag(eigen) %*% t(Q)

# generate data
Z = matrix(rnorm(10*50), nrow = 50, ncol = 10)
out = eigen(S, symmetric = TRUE)
S.sqrt = out$vectors %*% diag(out$values^0.5) %*% t(out$vectors)
X = Z %*% S.sqrt

```
<br>\vspace{0.5cm}

<!-- <center> -->
<!-- ![Figure caption here.](images/repsKLdense_N50_P10.png){ -->
<!-- width=90% } -->
<!-- </center> -->

```{r admmsim4, eval = T, echo = F, fig.cap = "Figure caption here."}

knitr::include_graphics("images/repsKLdense_N50_P10.png")

```

#### Tridiagonal: P = 100, N = 50

```{r, message = FALSE, echo = TRUE, eval = FALSE, tidy = FALSE}

# generate covariance matrix
# (can confirm inverse is tri-diagonal)
S = matrix(0, nrow = 100, ncol = 100)
for (i in 1:100){
  for (j in 1:100){
    S[i, j] = 0.7^abs(i - j)
  }
}
